Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

### 2023-10-26 manuscript/manuscript.qmd
Use Zotero + update all possible references

Over the last three decades, governments globally have increasingly sought to crack down on NGOs (CITE). Repressive governments have repressed these organizations using both violence, as well as administrative crackdown, which create barriers to funding and advocacy for a variety of NGOs. (CITE) Policymakers have referred to this phenomenon as the "closing of civic space" or "shrinking civic space" (CITE) and data from the CIVICUS Civil Society Monitor shows that civil society groups face serious restrictions in nearly 60% of countries—111 in total (UPDATE FROM NEW CIVICUS DATA). These attacks—most of which have been implemented through anti-NGO laws—have adversely affected the ability of both domestic and international NGOs to operate and raise funds. Official aid funds have decreased substantially in countries that have imposed restrictive anti-NGO legislation [@ChaudhryHeiss:2018; @DupuyPrakash:2018] and official aid donors channel fewer funds to contentious issues and instead choose to fund "tame" causes [@ChaudhryHeiss:2018; @Bush:2015, HERROLD, SPRINGMAN]. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to foreign individual donors to offset the loss of funding from government and foundation sources [@BanksHulmeEdwards:2015].

The amount of private foreign aid is not insignificant—from 2000–2010, private giving by U.S. entities towards international causes doubled in real terms [@desaikharas, 505]. In 2016, U.S. private philanthropy towards international NGOs amounted to \$22.06 billion [@givingusa:2017]. Most notably, a large portion of this growth was driven by individuals making small-scale donations. In 2017, individuals gave US\$286.7 billion or approximately 70% of total giving to international affairs. This further increased to US\$427.71 billion in 2018, 85.7% of which was through individuals (Giving USA, 2019). This turn to private donors opens INGOs up to new dynamics in seeking out aid. Many organizations that had previously relied on governments and foundations had optimized their fundraising strategies for those donors, but now need to tailor their fundraising strategies to private individuals, who behave differently from governments and foundations.

@@

These findings make two main contributions. While recent research on private foreign aid has examined funding decisions of donor NGOs and foundations[^1], we know less about the funding choices of the individuals responsible for much of these private aid flows [@desaikharas]. The systematic study of private foreign aid is still in its infancy—there is an absence of expansive and high quality data [@greenhill2013], and most research has been restricted to elite, high net-worth donors [@USTrust2014]. Our research makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individual donor preferences.

Second, research shows that NGO leades often have difficulty in attracting new donors and also find it challenging to moviate current donors (Cao 2015). Examining the organizational-level factors in a group that make the average donor increase their charitable-giving towards these groups, especially when facing restricted civic space abroad, can help NGOs better target their programming and frame their funding appeals towards these donors. Our results thus have important implications for international NGOs facing increased antagonism by government abroad and point to possible methods of adapting to worsening or shrinking civic environments.

@@

[^1]: On aid from world's largest NGOs, see @koch2009, on private aid flows from NGOs based in the US, see @buthe2012.

# What determines individual donor behvior?

Substantial research on NGOs has looked at the motivation of individual donors, including altruism, reputational benefits and alignments with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. However, this research has overwhelmingly looked at individual donor behavior and motivation for giving to local—rather than international—causes (CITE -2). Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor (Casale & Baumann, 2015, 100). Further, as Micklewright and Schnepf (2009) point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, “donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad” (Tremblay-Boire & Prakash, 2017, p. 644). However, we lack adequate data and theory regarding philanthropy towards INGOs.

The third wave of democratization, the collapse of the Soviet Union, and the spread of a "pro-NGO norm" led to an increased role and visibility for NGOs. Globally, a large amount of resources were channeled to NGOs focused on service provision or development because many public and private donors perceived them as more efficient than states with weak institutions and poor governance; there was a high risk that such states would use aid transfers inefficiently (Dietrich 2013).

But as INGOs became increasingly powerful actors in global policy circles, they faced a number of criticisms. Many organizations struggle with the conflicting imperatives of being large organizations that can operate efficiently at scale, while maintaining the grassroots contacts and consensus building that contributed to their early successes.\footnote{\cite{mitchell2020between}, 2; \cite{jalali2013, bano2012, rahman2006}.} Many nonprofit scandals involve some form of financial wrongdoing (Prakash and Gugerty 2010). Globally, around 5\% of revenue is lost annually due to fraud (Lamothe et al 2022). Misappropriation of assets and embezzlement schemes are among the most common forms of financial fraud (Lamthoe). These criticisms and concerns, in turn, have led donors to demand greater NGO responsiveness, especially in terms of financial transparency and accountability.

Donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming (Croson & Shang, 2011; Tremblay-Boire & Prakash, 2017). Nonprofits fundraising internationally may therefore rely on framing to control perceptions. Below, we theorize the effect of different frames surrounding a country’s legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

In theory, donors should scrutinize each aspect of an NGO’s organizational structure and programmatic performance prior to donation, but this rarely happens in practice (Tremblay-Boire & Prakash, 2017). Instead, individuals rely on heuristics to simplify their decision-making.

One common signal for donors is organizational efficacy or the belief that an individual's donation will make a real difference. This can manifest itself through multiple mechanisms such as the amount an organization spends on overhead costs, whether an organization makes its finances public, or whether an organization participates in voluntary accountability practices, such as listing organizational details on websites like GuideStar and Charity Navigator. An NGO's financial transparency, defined as the degree of completeness of information, provided by organizations to the public concerning its business activities, is one tangible heuristics that donors use to make their decisions. NGOs can engage in financial transparency in a variety of ways such as distribution of audiented financial statements, third party intermediaries. Reserach shows that transparency is associated with greater donations, lower debt and better governance (Harris and Neely 2021, Saxton et al 2012). It can also enhance donor trust and confidence in organizations [@Vacarro:2006; @vaccaromadsen].

Previous research has found that media exposés about charity mismanagement can generate negative reputational spillovers for the charitable sector [@gibelmangelman]. Diversions, or the unauthorized use of an organization's assets, including embezzlement and theft can result in a decrease in donatons. This effect becomes even stronger with media coverage of diversion (Harris et al 2023). Organizational financial transparent can reduce actual or perceived information asymmetries between donors and charities, this potentially increasing donor likelihood of donating to an INGO.

@@

Accountability is another factor that can impact donor willingness to donate. Accountability is necessary to promote public trust in NGOs. While transparency is an organization's willingness to make critical data about the organization public, accountability is the willingness of the organization to explain its action to shareholders. It also encompasses how well an organization's mission is aligned with its reources and how it manages resources to optimize performance (Sloan 2009).

Explaining actions to stakeholders can be important because many NGOs, even when well-intentioned ones, can devote excessive resources towards marketing and overheads [@bowman:2006]. Private donors, subsequently, may be more likely to dislike NGOs that spend a large share of their revenues for unproductive purposes [@rose1982], or rather, a large share of resources on administration, management and fundraising. As @Tremblay-BoirePrakash:2017 point out, while individuals would like to have some assurance that their resources will be used appropriately and organizations are not spending too much on overhead, donors do not necessarily have the time to research charities thoroughly. Lack of accountability is associated with lesser public trust, reputation, perceived quality and donation behavior (Becker 2018).

To judge organizational accountabiltiy, donors may use ratings by watchdog organizations like Charity Navigator and GuideStar as heuristics to guide their decisions. Research shows that initiatives to address accountability gaps by these organizations can affect donations to organizations in a significant manner (Sloan 2009, Gordon, Knock and Neely).

@@

In addition to organizational factors discussed above, political contexts in NGO host countries, especially their environments for civil society organizations, can also influence donor preferences for donating to foreign NGOs. A global cascade of anti-NGO laws in recent decades has created barriers to entry, funding, and advocacy for civic organizations (Glasius et al, Schmidt et al, Chaudhry, Heiss, DPR). As a result, NGOs must spend more time, effort, and resources to ensure their survival -- at the expense of pursuing their missions. Some NGOs have adapted to their repressive environments by recreating their organizations and changing their mission to avoid directly confronting the government. Analyzing the impact of anti-NGO laws in Bangladesh and Zambia, scholars have similarly found that NGOs with a broad mission ended up restricting themselves to service work and giving up on their advocacy work. Those that continued to engage in transnational advocacy changed the targets of their advocacy, the issues they focused on, and the language they used. For instance, many NGOs stopped lobbying the central government and focused their advocacy efforts on lower levels of government.\footnote{\cite{fransen2021tempering}.} Similarly, in Russia, many NGOs working with foreign partners and lobbying the central government no longer consider it an effective strategy.\footnote{\cite{sundstrom2022evolution}.} Overall, fewer groups may end up focusing on contentious causes, instead focusing on regime-compatible causes.

The changes in civic environments impacting NGO activity have also influenced official donor responses. In response to the passage of these laws, scholars have found that while multilateral donors do not reduce aid following the passage of anti-NGO laws, they are associated with a 32\% decline in bilateral aid inflows.\footnote{\cite{DupuyPrakash:2018}.} Further, advocacy-oriented donors -- i.e., those that fund democracy and civil society promotion activities -- reduced their spending by 74\%, but did not cut their spending on development projects such as education, health, water, and sanitation.\footnote{\cite{right2022advocacy}. However, previous literature has established that individual private donors may not necessarily use the same metrics as official donors (CITE). Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@easterlywilliamson]. These anti-NGO laws may signal to private donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization.

@@

Identifying the causal link between structural factors and donor behavior is more difficult than measuring the link between organizational factors and donor behavior, however, because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like financial transparency and accountability can sour the NGO-government relationship -- governments may be more likely to target NGOs that lack financial transparency and accountability. @fig:big-dag helps disentangle this complex relationship. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both transparency and accountability. As such, legal crackdowns do not occur in isolation. Countries are more likely to repress NGOs when organizational features are in tension with government preferences and when organizations receive substantial funding from abroad [@DPR2015]. Therefore, the same organizational characteristic that can increase the likelihood of crackdown can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

@@

Since we are running a conjoint experiment, the interactions between the main treatment variables described above is simply an exploratory analysis. Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Moreover, the average donor will not know exactly what “crackdown” entails for any given organization (i.e., does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified short- cuts to donor information-gathering. Knowing that an organization is financially transparent (even if donors are not familiar with specific kinds of requirements regarding transparency) or knowing that an NGO faces legal trouble abroad should be enough to trigger the effect of these frames on donor preferences. Moreover, although many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, transparency and accountability. Instead, donors rely on signals and heuristics about the organiza- tion to shape their preferences and behavior (Szper & Prakash, 2011).

@@

# Conclusion
Over the last two decades, governments globally have systematically controlled and repressed NGOs using both violent tactics and administrative means over the last three decades -- a phenomenon known as `closing' or `shrinking' civic space. This trend of closing civic space has important implications for local and international NGOs as well as donors. Recent research has analyzed the impact of this crackdown on public or official aid donors. However, we know relatively little about how state repression of NGOs affects the preferences of foreign private donors. The responses of elite donors and foundations to government crackdown on civil society organizations remains a ripe area for research. Unlike state donors, individual and elite donors may not necessarily withdraw or reduce support for NGOs facing harassment abroad. This is important because private philanthropy towards organizations working in international affairs continues to grow despite inflation -- from 2021 to 2022, individual giving to nonprofits in international affairs grew by over 10.9% in the US (https://philanthropy.iupui.edu/news-events/news-item/giving-usa:-total-u.s.-charitable-giving-declined-in-2022-to-$499.33-billion-following-two-years-of-record-generosity.html?id=422).

@@

These results have important implications for NGO operations, fundraising and survivability at a time when a number of INGOs are dealing with hostile host governments. The crackdown is even more challenging for local NGOs in Global South countries, where much government and public support is channeled via service delivery organizations. Local groups tackling contentious issues may not be able to raise funds domestically, as publics may prefer donating to service organizations focusing on health, education, and poverty.\footnote{\cite{absar2017will}.} Even local philanthropists may be deterred from giving to organizations focusing on contentious causes such as advocacy, media freedom, and anti-corruption initiatives due to poor tax incentives or a fear of retribution.\footnote{\cite{Brechenmacher:2017}; \cite{DPR2015}; \cite{Baoumi:2016}.} Our results highlight the importance of organizational characteristics like transparency and accountability even in an era of closing civic space. By finding new ways to communicate about organizational transparency and accountability to potential donors, NGOs can help overcome individual donors' concerns about host governments' criticisms of and crackdown on these groups.

>>>>

Over the last three decades, governments globally have increasingly sought to crack down on NGOs (CITE). Repressive governments have repressed these organizations using both violence, as well as administrative crackdown, which create barriers to funding and advocacy for a variety of NGOs. (CITE) Policymakers have referred to this phenomenon as the "closing of civic space" or "shrinking civic space" (CITE) and data from the CIVICUS Civil Society Monitor shows that civil society groups face serious restrictions in nearly 60% of countries—111 in total (UPDATE FROM NEW CIVICUS DATA). These attacks—most of which have been implemented through anti-NGO laws—have adversely affected the ability of both domestic and international NGOs to operate and raise funds. Official aid funds have decreased substantially in countries that have imposed restrictive anti-NGO legislation [@ChaudhryHeiss:2018; @DupuyPrakash:2018] and official aid donors channel fewer funds to contentious issues and instead choose to fund "tame" causes [@ChaudhryHeiss:2018; @Bush:2015; @HERROLD; @SPRINGMAN]. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to foreign individual donors to offset the loss of funding from government and foundation sources [@BanksHulmeEdwards:2015].

The amount of private foreign aid is not insignificant—from 2000–2010, private giving by U.S. entities towards international causes doubled in real terms [@DesaiKharas:2018, 505]. In 2016, U.S. private philanthropy towards international NGOs amounted to \$22.06 billion [@GivingUSA:2017]. Most notably, a large portion of this growth was driven by individuals making small-scale donations. In 2017, individuals gave US\$286.7 billion or approximately 70% of total giving to international affairs. This further increased to US\$427.71 billion in 2018, 85.7% of which was through individuals [@GivingUSA:2019]. This turn to private donors opens INGOs up to new dynamics in seeking out aid. Many organizations that had previously relied on governments and foundations had optimized their fundraising strategies for those donors, but now need to tailor their fundraising strategies to private individuals, who behave differently from governments and foundations.

@@

These findings make two main contributions. While recent research on private foreign aid has examined funding decisions of donor NGOs and foundations[^1], we know less about the funding choices of the individuals responsible for much of these private aid flows [@DesaiKharas:2018]. The systematic study of private foreign aid is still in its infancy—there is an absence of expansive and high quality data [@GreenhillPrizzonRogerson:2013], and most research has been restricted to elite, high net-worth donors [@USTrust2014]. Our research makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individual donor preferences.

Second, research shows that NGO leaders often have difficulty in attracting new donors and also find it challenging to motivate current donors [@GaudreauCao:2015]. Examining the organizational-level factors in a group that make the average donor increase their charitable-giving towards these groups, especially when facing restricted civic space abroad, can help NGOs better target their programming and frame their funding appeals towards these donors. Our results thus have important implications for international NGOs facing increased antagonism by government abroad and point to possible methods of adapting to worsening or shrinking civic environments.

@@

[^1]: On aid from world's largest NGOs, see @Koch:2009, on private aid flows from NGOs based in the US, see @ButheMajorMello-e-Souza:2012.

# What determines individual donor behavior?

Substantial research on NGOs has looked at the motivation of individual donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. However, this research has overwhelmingly looked at individual donor behavior and motivation for giving to local—rather than international—causes (CITE -2). Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100]. Further, as @MicklewrightSchnepf:2009 point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, “donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad” [@Tremblay-BoirePrakash:2017, 644]. However, we lack adequate data and theory regarding philanthropy towards INGOs.

The third wave of democratization, the collapse of the Soviet Union, and the spread of a "pro-NGO norm" led to an increased role and visibility for NGOs. Globally, a large amount of resources were channeled to NGOs focused on service provision or development because many public and private donors perceived them as more efficient than states with weak institutions and poor governance; there was a high risk that such states would use aid transfers inefficiently [@Dietrich:2013].

But as INGOs became increasingly powerful actors in global policy circles, they faced a number of criticisms. Many organizations struggle with the conflicting imperatives of being large organizations that can operate efficiently at scale, while maintaining the grassroots contacts and consensus building that contributed to their early successes [@MitchellSchmitzBruno-vanVijfeijken:2020; @Jalali:2008; @Bano:2012; @Rahman:2006]. Many nonprofit scandals involve some form of financial wrongdoing [@PrakashGugerty:2010b]. Globally, around 5% of revenue is lost annually due to fraud [@ACFE:2021]. Misappropriation of assets and embezzlement schemes are among the most common forms of financial fraud [@LamotheTer-MkrtchyanRuddle:2023]. These criticisms and concerns, in turn, have led donors to demand greater NGO responsiveness, especially in terms of financial transparency and accountability.

Donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. Nonprofits fundraising internationally may therefore rely on framing to control perceptions. Below, we theorize the effect of different frames surrounding a country’s legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

In theory, donors should scrutinize each aspect of an NGO’s organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, individuals rely on heuristics to simplify their decision-making.

One common signal for donors is organizational efficacy or the belief that an individual's donation will make a real difference. This can manifest itself through multiple mechanisms such as the amount an organization spends on overhead costs, whether an organization makes its finances public, or whether an organization participates in voluntary accountability practices, such as listing organizational details on websites like GuideStar and Charity Navigator. An NGO's financial transparency, defined as the degree of completeness of information, provided by organizations to the public concerning its business activities, is one tangible heuristics that donors use to make their decisions. NGOs can engage in financial transparency in a variety of ways such as distribution of audited financial statements, third party intermediaries. Research shows that transparency is associated with greater donations, lower debt and better governance [@HarrisNeely:2021; @SaxtonKuoHo:2012]. It can also enhance donor trust and confidence in organizations [@Vaccaro:2006; @VaccaroMadsen:2006].

Previous research has found that media exposés about charity mismanagement can generate negative reputational spillovers for the charitable sector [@GibelmanGelman:2004]. Diversions, or the unauthorized use of an organization's assets, including embezzlement and theft can result in a decrease in donations. This effect becomes even stronger with media coverage of diversion [@HarrisPetrovitsYetman:2023]. Organizational financial transparent can reduce actual or perceived information asymmetries between donors and charities, this potentially increasing donor likelihood of donating to an INGO.

@@

Accountability is another factor that can impact donor willingness to donate. Accountability is necessary to promote public trust in NGOs. While transparency is an organization's willingness to make critical data about the organization public, accountability is the willingness of the organization to explain its action to shareholders. It also encompasses how well an organization's mission is aligned with its resources and how it manages resources to optimize performance [@Sloan:2009].

Explaining actions to stakeholders can be important because many NGOs, even when well-intentioned ones, can devote excessive resources towards marketing and overheads [@Bowman:2006]. Private donors, subsequently, may be more likely to dislike NGOs that spend a large share of their revenues for unproductive purposes [@Rose-Ackerman:1982], or rather, a large share of resources on administration, management and fundraising. As @Tremblay-BoirePrakash:2017 point out, while individuals would like to have some assurance that their resources will be used appropriately and organizations are not spending too much on overhead, donors do not necessarily have the time to research charities thoroughly. Lack of accountability is associated with lesser public trust, reputation, perceived quality and donation behavior [@Becker:2018].

To judge organizational accountability, donors may use ratings by watchdog organizations like Charity Navigator and GuideStar as heuristics to guide their decisions. Research shows that initiatives to address accountability gaps by these organizations can affect donations to organizations in a significant manner [@Sloan:2009; @GordonKnockNeely:2009]

@@

In addition to organizational factors discussed above, political contexts in NGO host countries, especially their environments for civil society organizations, can also influence donor preferences for donating to foreign NGOs. A global cascade of anti-NGO laws in recent decades has created barriers to entry, funding, and advocacy for civic organizations [@GlasiusSchalkDeLange:2020; @Schmidt-et-al; @Chaudhry:2022; @Heiss:2017; @DupuyRonPrakash:2015]. As a result, NGOs must spend more time, effort, and resources to ensure their survival -- at the expense of pursuing their missions. Some NGOs have adapted to their repressive environments by recreating their organizations and changing their mission to avoid directly confronting the government. Analyzing the impact of anti-NGO laws in Bangladesh and Zambia, scholars have similarly found that NGOs with a broad mission ended up restricting themselves to service work and giving up on their advocacy work. Those that continued to engage in transnational advocacy changed the targets of their advocacy, the issues they focused on, and the language they used. For instance, many NGOs stopped lobbying the central government and focused their advocacy efforts on lower levels of government [@FransenDupuyHinfelaar:2021]. Similarly, in Russia, many NGOs working with foreign partners and lobbying the central government no longer consider it an effective strategy [@SundstromHenrySperling:2022]. Overall, fewer groups may end up focusing on contentious causes, instead focusing on regime-compatible causes.

The changes in civic environments impacting NGO activity have also influenced official donor responses. In response to the passage of these laws, scholars have found that while multilateral donors do not reduce aid following the passage of anti-NGO laws, they are associated with a 32% decline in bilateral aid inflows [@DupuyPrakash:2018]. Further, advocacy-oriented donors -- i.e., those that fund democracy and civil society promotion activities -- reduced their spending by 74%, but did not cut their spending on development projects such as education, health, water, and sanitation [@RightSpringmanWibbels:2022]. However, previous literature has established that individual private donors may not necessarily use the same metrics as official donors (CITE). Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. These anti-NGO laws may signal to private donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization.

@@

Identifying the causal link between structural factors and donor behavior is more difficult than measuring the link between organizational factors and donor behavior, however, because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like financial transparency and accountability can sour the NGO-government relationship -- governments may be more likely to target NGOs that lack financial transparency and accountability. @fig:big-dag helps disentangle this complex relationship. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both transparency and accountability. As such, legal crackdowns do not occur in isolation. Countries are more likely to repress NGOs when organizational features are in tension with government preferences and when organizations receive substantial funding from abroad [@DupuyRonPrakash:2015]. Therefore, the same organizational characteristic that can increase the likelihood of crackdown can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

@@

Since we are running a conjoint experiment, the interactions between the main treatment variables described above is simply an exploratory analysis. Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Moreover, the average donor will not know exactly what “crackdown” entails for any given organization (i.e., does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified short- cuts to donor information-gathering. Knowing that an organization is financially transparent (even if donors are not familiar with specific kinds of requirements regarding transparency) or knowing that an NGO faces legal trouble abroad should be enough to trigger the effect of these frames on donor preferences. Moreover, although many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, transparency and accountability. Instead, donors rely on signals and heuristics about the organization to shape their preferences and behavior [@SzperPrakash:2011].

@@

# Conclusion
Over the last two decades, governments globally have systematically controlled and repressed NGOs using both violent tactics and administrative means over the last three decades -- a phenomenon known as `closing' or `shrinking' civic space. This trend of closing civic space has important implications for local and international NGOs as well as donors. Recent research has analyzed the impact of this crackdown on public or official aid donors. However, we know relatively little about how state repression of NGOs affects the preferences of foreign private donors. The responses of elite donors and foundations to government crackdown on civil society organizations remains a ripe area for research. Unlike state donors, individual and elite donors may not necessarily withdraw or reduce support for NGOs facing harassment abroad. This is important because private philanthropy towards organizations working in international affairs continues to grow despite inflation -- from 2021 to 2022, individual giving to nonprofits in international affairs grew by over 10.9% in the US [@GivingUSA:2023].

@@

These results have important implications for NGO operations, fundraising and survivability at a time when a number of INGOs are dealing with hostile host governments. The crackdown is even more challenging for local NGOs in Global South countries, where much government and public support is channeled via service delivery organizations. Local groups tackling contentious issues may not be able to raise funds domestically, as publics may prefer donating to service organizations focusing on health, education, and poverty [@absar2017will]. Even local philanthropists may be deterred from giving to organizations focusing on contentious causes such as advocacy, media freedom, and anti-corruption initiatives due to poor tax incentives or a fear of retribution [@Brechenmacher:2017; @DupuyRonPrakash:2015; @Baoumi:2016]. Our results highlight the importance of organizational characteristics like transparency and accountability even in an era of closing civic space. By finding new ways to communicate about organizational transparency and accountability to potential donors, NGOs can help overcome individual donors' concerns about host governments' criticisms of and crackdown on these groups.
